---
title: "Test_r"
output: pdf_document
date: "2025-10-29"
---


```{r load-data, include=FALSE}
library(readr)
pokemon_df <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-04-01/pokemon_df.csv")
```

```{r libraries, message=FALSE}
library(tidyverse)
```
# Introduction

The chosen dataset comes from tidytuesday and is a dataset regarding Pokemons. It can be found here: https://github.com/rfordatascience/tidytuesday/blob/main/data/2025/2025-04-01


# Data presentation

A summery of the data set can be seen below. From this information, it is possible to see which variables makes sense to do statistical test's on. 

As an example "id" and "species_id" does not make sense to use in these test's, because it is a index value more that data value. 

Therefore for this assignment, the chosen variables for further statistical investigation are the Pokémons "attack" and "hp" data. 

```{r Summery}
summary(pokemon_df)
```
To visualize the data, one can plot the two variables against each other. As seen below:

```{r pokemon-plot, fig.width=7, fig.height=5}
ggplot(pokemon_df, aes(x = attack , y = hp)) +
  geom_point(alpha = 0.7, size = 3) +
  theme_minimal() +
  labs(
    title = "Pokémon: Attack vs HP",
    x = "Attack",
    y = "HP"
  )
```

We setup a test, that tests if there is a correlation between the two variables. 

Here our null hypothesis is that the more HP a pokémon has the more attack it will have as well
```{r test_stat}
cor_result <- cor.test(pokemon_df$attack, pokemon_df$hp)

cor_result$estimate  # correlation coefficient
cor_result$p.value   # p-value


linear_model <- lm(attack ~ hp, data=pokemon_df)
summary(linear_model)

```

From this we have a very low p-value, meaning that we will have to reject our null hypothesis. 
Furthermore if one where to fit a linear model to this data we can see the model fit very poorly, both visually as seen below, but it is also possible to examine the correlation coefficient of the fit. 


```{r}
library(ggplot2)

ggplot(pokemon_df, aes(x = hp, y = attack)) +
  geom_point(alpha = 0.6, color = "darkblue") +       # scatter points
  geom_smooth(method = "lm", se = TRUE, color = "red") +  # regression line
  theme_minimal() +
  labs(
    title = "Attack vs HP of Pokémon",
    subtitle = paste0("Correlation: ", round(cor_result$estimate, 2),
                      ", p-value: ", signif(cor_result$p.value, 3)),
    x = "HP",
    y = "Attack"
  )

```

To make the next test a combination of the offensive and defensive stats respectively of the pokémon have been made.
Considering attack, special_attack and speed offensive traits, and defense, special_defense and hp to be defensive traits.
The combination of these is made with a simple summation, and save in new columns called offensive and defensive.
This is done based on the hypothesis that a pokémon with with combined higher offensive stats would have lower defensive stats and vise versa.


```{r bind data}

pokemon_df <- pokemon_df %>%
  mutate(offensive = attack + special_attack + speed)

pokemon_df <- pokemon_df %>%
  mutate(defensive = defense + special_defense + hp)
```

Using the linear model like before we test to see if there is a correlation between combined offensive and defensive stats of a pokémon.
Again we get a p-value in the 2e-16 range which means we see no statistical correlation between the offensive and defensive stats.



```{r}
linear_model_off_deff <- lm(offensive ~ defensive, data=pokemon_df)
summary(linear_model_off_deff)
```
To visualize the same code is reused.

```{r plot offensive/defensive}
cor_result <- cor.test(pokemon_df$defensive, pokemon_df$offensive)

cor_result$estimate  # correlation coefficient
cor_result$p.value   # p-value

ggplot(pokemon_df, aes(x = defensive, y = offensive)) +
  geom_point(alpha = 0.6, color = "darkblue") +       # scatter points
  geom_smooth(method = "lm", se = TRUE, color = "red") +  # regression line
  theme_minimal() +
  labs(
    title = "Combined offensive vs defensive stats of Pokémon",
    subtitle = paste0("Correlation: ", round(cor_result$estimate, 2),
                      ", p-value: ", signif(cor_result$p.value, 3)),
    x = "HP",
    y = "Attack"
  )
```







```
